---
title: "Carbon Inventory Code - Prepared in R Studio"
author: "Carbon Counters"
date: "3/18/2021"
output: word_document
---

This code was used to prepare the full landscape carbon inventory, as well as the working lands inventories used in the baseline reference scenario code (see supplemental materials). For all data citations and detailed explanation of methods, please refer to the Methodology section of the report.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache.lazy = FALSE)
```

## Code setup

```{r}

# Attach packages
library(tidyverse)
library(tidyr)
library(here)
library(janitor)
library(plotly)
library(kableExtra)
library(effsize)
library(broom)
library(formattable)
library(purrr)

# read in data. file descriptions above each file.

# carbon value table from CARB
carbon_vals <- read.csv(here::here("files", "luts", "lut_lf_carb.csv"), encoding = "UTF-8")

# existing vegetation cover (LANDFIRE) look up table
evc_lut <- read.csv(here::here("files", "luts", "lut_evc.csv"), encoding = "UTF-8") %>%
  clean_names() %>%
  rename(classnames_evc = x_u_feff_classnames)

# existing vegetation height (LANDFIRE) look up table
evh_lut <- read.csv(here::here("files", "luts", "lut_evh.csv"), encoding = "UTF-8") %>% 
  clean_names() %>% 
  rename(classnames_evh = x_u_feff_classnames)

# existing vegetation type (LANDFIRE) look up table
evt_lut <- read.csv(here::here("files", "luts", "lut_evt.csv"), encoding = "UTF-8") %>% 
  clean_names() %>% 
  rename(classnames_evt = x_u_feff_evt_name)

# nitrogen look up table (2007 SB County GHG inventory)
lut_n <-  read.csv(here::here("files", "luts", "lut_n.csv"), encoding = "UTF-8") %>% 
  clean_names() %>% 
  rename(nitrogen_cat = x_u_feff_nitrogen)

# 2016 evc/evh/evt for 2016, assigned to points for each 30x30m pixel in the county (our pre-processed data)
lf_evc_16 <- read_csv(here::here("files", "natlands", "LF_2016_EVC.csv"))

lf_evh_16 <- read_csv(here::here("files",  "natlands", "LF_2016_EVH.csv"))

lf_evt_16 <- read_csv(here::here("files",  "natlands", "LF_2016_EVT.csv"))

# read in nitrogen categories to apply to LANDFIRE data where appropriate
lf_reclass_n <- read.csv(here::here("files", "luts", "lf_reclass_nitrogen.csv"), encoding = "UTF-8") %>% 
  clean_names() %>% 
  rename(classnames_evt = x_u_feff_evt) %>% 
  rename(lf_n_category = n_category)

# read in soil data - unit = gC / m^2

soil <- read_csv(here::here("files", "soil", "ssurgo.csv")) %>% 
  dplyr::select(pointid, soc0_30)

```


## 2016 full landscape inventory

```{r}

###############################################################
# Natural lands
###############################################################

# Data cleaning and merging

carbon_vals <- carbon_vals %>% 
  clean_names("snake") %>% 
  filter(!is.na(w_total_mt_cha)) %>% 
  rename(total_mt = w_total_mt_cha) %>% 
  rename(grouped = x_u_feff_lf_key)

# combine all three (EVC, EVH, EVT) landfire dataframes for 2016, and clean
precombin_df <- merge(lf_evc_16, lf_evh_16, by = "OBJECTID")

# combine all together, and create "grouped" variable to later match with carbon values
combined_lf_df <- merge(precombin_df, lf_evt_16, by = "OBJECTID") %>% 
  dplyr::select(OBJECTID, pointid, CLASSNAMES.x, CLASSNAMES.y, EVT_NAME,  Reclass_16) %>% 
  clean_names("snake") %>% 
  rename(classnames_evc = classnames_x) %>% 
  rename(classnames_evh = classnames_y) %>%
  rename(classnames_evt = evt_name) %>% 
  left_join(evh_lut, by = "classnames_evh") %>% 
  left_join(evc_lut, by = "classnames_evc") %>% 
  left_join(evt_lut, by = "classnames_evt") %>% 
  mutate(grouped = paste(evt_group, evh_group, evc_group, sep = "")) %>% 
  left_join(lf_reclass_n, by = "classnames_evt") %>% 
  dplyr::select(-reclass_16) %>% 
  rename(reclass_16 = reclass_category)

###############################################################
# Add agricultural lands
###############################################################

# read in ag data

ag_2016_raw <- read.csv(here::here("files", "ag", "ag_2016.csv"), encoding = "UTF-8", na.strings=c(""," ", "NoData", "NA"))

# clean 2016 ag data and make appropriate changes to nitrogen values and classifications (some of these may be outdated/irrelevant, but this does not affect the results)
# designate which points will rely on CalAg data, and which on LANDFIRE. This effectively overlays the CalAg data onto the LANDFIRE data

ag_2016 <- ag_2016_raw %>% 
  dplyr::select(!c(organic, crop_list)) %>% 
  rename(nitrogen = nitrogren_) %>% 
  clean_names("snake") %>% 
  rename(pointid = objectid) %>% 
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(nitrogen = as.character(nitrogen)) %>% 
  mutate(nitrogen = ifelse(ag_class == "Barren / Fallow" | ag_class == "Greenhouse", 0, nitrogen)) %>% 
  mutate(ag_class = ifelse(ag_class == "Irrigated Pasture",  "Fodder", as.character(ag_class))) %>% 
  mutate(ag_class = ifelse(ag_class == "Barren / Fallow",  "Fallow", as.character(ag_class))) %>% 
  mutate(nitrogen = ifelse(ag_class == "Pastureland", "Field Crops", nitrogen)) %>% 
  mutate(source = ifelse(is.na(ag_class), "landfire", "calag"))

# merge natural lands with ag, replace grouped name with ag classification where appropriate, clean data

combined_ag_natland <- merge(combined_lf_df, ag_2016, by = "pointid") %>%
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(lf_n_category = as.character(lf_n_category)) %>% 
  mutate(grouped = ifelse(is.na(ag_class), grouped, ag_class)) %>% 
  mutate(reclass_cat = ifelse(is.na(ag_class), as.character(reclass_16), as.character(ag_class))) %>%
  mutate(reclass_cat = as.character(reclass_cat)) %>% 
  mutate(reclass_cat = ifelse(reclass_cat == "Barren / Fallow",  "Fallow", as.character(reclass_cat))) %>% 
  dplyr::select(evt_group, pointid, reclass_cat, grouped, nitrogen, lf_n_category, source) %>% 
  rename(nitrogen_cat = nitrogen) %>% 
  mutate(reclass_cat = ifelse(reclass_cat == "Wetland",  "Riparian/Wetland", as.character(reclass_cat))) %>%
  mutate(reclass_cat = ifelse(reclass_cat == "Irrigated Pasture",  "Fodder", as.character(reclass_cat))) %>%
  mutate(nitrogen_cat = ifelse(source == "calag", as.character(nitrogen_cat), as.character(lf_n_category)))

# create simplified file to use in GIS mapping
reclass_map_file <- combined_ag_natland %>% 
  dplyr::select(pointid, reclass_cat)

# calculate stored carbon and nitrous oxide emissions for each pixel (900 sq m)

ag_natland_carbon_n_16 <- combined_ag_natland %>% 
  left_join(carbon_vals, by = "grouped") %>%
  mutate(mt_900 = (total_mt*.09)) %>% # MT carbon per hectare multiplied by .09 to get metric tons of carbon per pixel (900 sq m)
  left_join(lut_n, by = "nitrogen_cat") %>% 
  mutate(lbs_n_pixel = (n_rate_lbs_acre*.222395)) %>% # nitrogen application rate (pounds per acre) multiplied by .222395 to get pounds of N applied per per pixel
  mutate(emit_n_lbs_pix = (lbs_n_pixel * .0175)) %>% # 1.75% of nitrogen escapes at N2O emissions
  dplyr::select(!c(n_rate_lbs_acre, lbs_n_pixel)) %>% 
  mutate(stock_abvgc_mtco2e_pixel = (mt_900*1)) %>%  # multiply metric tons of carbon by 3.67 to get MT of CO2 equivalent # Decided to report as MT instead, replace 3.67 value w 1 to not break rest of code
  mutate(emit_no_mtco2e_pix = emit_n_lbs_pix*298*0.000453592) # multiply pounds to N2O emissions by 298 to convert to pounds CO2e, then by 0.000453592 to get metric tonnes

# combine aboveground carbon and nitrogen data with soil data
         
all_c_n_soil <- merge(ag_natland_carbon_n_16, soil, by = "pointid") %>% 
  mutate(soil900 = (soc0_30*900)) %>% #per m^2 to per 900 m^2
  mutate(soilMT = (soil900/1000000)) %>%  # grams to metric tons of organic carbon
  mutate(stock_soilc_mtco2e_pix = soilMT*1) # convert to CO2e # Decided to report as MT instead, replace 3.67 value w 1 to not break rest of code
  
# make into a table, does not yet include urban forestry

all_clean_16_no_tree <- all_c_n_soil %>% 
  dplyr::select(pointid, reclass_cat, stock_abvgc_mtco2e_pixel, stock_soilc_mtco2e_pix, emit_no_mtco2e_pix, source) %>% 
  mutate(emit_no_mtco2e_pix = replace_na(emit_no_mtco2e_pix, 0))

# calculate acreages per landclass category

all_acreages_16 <- all_clean_16_no_tree %>% 
  group_by(reclass_cat) %>% 
  summarize(pixels = n()) %>% 
  mutate(sqmeter = pixels*900) %>% 
  mutate(acreage = sqmeter/4047) %>% 
  dplyr::select(! c(pixels, sqmeter)) %>% 
  adorn_totals()

#create a summary (preliminary inventory without urban forestry)
ci_summary_cat_16 <- all_clean_16_no_tree %>%
  dplyr::select(!source) %>% 
  group_by(reclass_cat) %>%
  summarise_all(.funs = c(sum="sum"), na.rm = TRUE) %>%
  mutate(net = (stock_soilc_mtco2e_pix_sum + stock_abvgc_mtco2e_pixel_sum)) %>%
  merge(all_acreages_16, by = "reclass_cat") %>% 
  dplyr::select(!pointid_sum)

###############################################################
# Add in urban forestry
###############################################################

# CO2e stored/urban tree canopy (metric tons/acre) converted to MT C

tree_num <- 114.8730627/3.67

# Assign # of acres of developed land as object for calculations
# If making changes, double check this cell reference is correct (should reference "developed")

urban_acres <- all_acreages_16[2, 2]

# create a row that matches ci_summary_cat_16 (preliminary inventory) to add data with rbind
 
tree_row <-data.frame("Urban Forestry (Aboveground Only)", tree_num*urban_acres, 0, 0, tree_num*urban_acres, 0)
names(tree_row)<-c("reclass_cat", "stock_abvgc_mtco2e_pixel_sum", "stock_soilc_mtco2e_pix_sum", "emit_no_mtco2e_pix_sum", "net", "acreage")

# add urban forestry row to inventory

ci_summary_cat_16 <- rbind(ci_summary_cat_16, tree_row) %>% 
  mutate(reclass_cat = ifelse(reclass_cat == "Urban Forestry (Aboveground Only)", "Developed", reclass_cat)) %>% 
  group_by(reclass_cat) %>% 
  summarise_all(sum) %>% 
  adorn_totals()

# rename columns

colnames(ci_summary_cat_16) = c("Landcover Classification", "Total Aboveground Carbon (MT C)", "Total Soil Carbon (MT C)", "Total NO Emissions (MT CO2e)", "Total Stocks (MT C)", "Acres")

# final landscape inventory
ci_summary_cat_16
```

The next section essentially repeats the process above, using a loop, but excludes LANDFIRE data to create inventories from the CalAg data for 2012, 2016, and 2019. These will be used to create the baseline reference scenario

```{r}

###############################################################
# Process dataframes to use in baseline projections
###############################################################

# read in ag 2012 and 2019,and re-read in 2016 data

ag_2019_raw <- read.csv(here::here("files", "ag", "ag_2019.csv"), encoding = "UTF-8", na.strings=c(""," ", "NoData", "NA"))

ag_2019 <- ag_2019_raw %>% 
  dplyr::select(!c(organic, crop_list)) %>% 
  rename(nitrogen = nitrogren_) %>% 
  clean_names("snake") %>% 
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(nitrogen = as.character(nitrogen)) %>% 
  mutate(nitrogen = ifelse(ag_class == "Barren / Fallow" | ag_class == "Greenhouse", 0, nitrogen)) %>% 
  mutate(ag_class = ifelse(ag_class == "Irrigated Pasture",  "Fodder", as.character(ag_class))) %>% 
  mutate(ag_class = ifelse(ag_class == "Barren / Fallow",  "Fallow", as.character(ag_class))) %>% 
  rename(pointid = objectid) %>% 
  mutate(source = ifelse(is.na(ag_class), "landfire", "calag"))

ag_2012_raw <- read.csv(here::here("files", "ag", "ag_2012.csv"), encoding = "UTF-8", na.strings=c(""," ", "NoData", "NA"))

ag_2012 <- ag_2012_raw %>% 
  dplyr::select(!c(organic, crop_list)) %>% 
  rename(nitrogen = nitrogren_) %>% 
  clean_names("snake") %>% 
  rename(pointid = objectid) %>% 
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(nitrogen = as.character(nitrogen)) %>% 
  mutate(nitrogen = ifelse(ag_class == "Barren / Fallow" | ag_class == "Greenhouse", 0, nitrogen)) %>% 
  mutate(ag_class = ifelse(ag_class == "Irrigated Pasture",  "Fodder", as.character(ag_class))) %>% 
  mutate(ag_class = ifelse(ag_class == "Barren / Fallow",  "Fallow", as.character(ag_class))) %>% 
  mutate(nitrogen = ifelse(ag_class == "Pastureland", "Field Crops", nitrogen)) %>% 
  mutate(source = ifelse(is.na(ag_class), "landfire", "calag"))

ag_2016_update <- ag_2016_raw %>% 
  dplyr::select(!c(organic, crop_list)) %>% 
  rename(nitrogen = nitrogren_) %>% 
  clean_names("snake") %>% 
  rename(pointid = objectid) %>% 
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(nitrogen = as.character(nitrogen)) %>% 
  mutate(nitrogen = ifelse(ag_class == "Barren / Fallow" | ag_class == "Greenhouse", 0, nitrogen)) %>% 
  mutate(ag_class = ifelse(ag_class == "Irrigated Pasture",  "Fodder", as.character(ag_class))) %>% 
  mutate(ag_class = ifelse(ag_class == "Barren / Fallow",  "Fallow", as.character(ag_class))) %>% 
  mutate(nitrogen = ifelse(ag_class == "Pastureland", "Field Crops", nitrogen)) %>% 
  mutate(source = ifelse(is.na(ag_class), "landfire", "calag"))

# put files into a list

ag_files_list <- list(ag_2012, ag_2016_update, ag_2019)

# first, merge with LANDFIRE data so as to not break any of the above code

fx_merge <- function(ag) {
  
  merge(combined_lf_df, ag, by = "pointid") %>%
  mutate(ag_class = as.character(ag_class)) %>% 
  mutate(grouped = ifelse(is.na(ag_class), grouped, ag_class)) %>% 
  mutate(reclass_cat = ifelse(is.na(ag_class), reclass_16, ag_class)) %>% 
  dplyr::select(evt_group, pointid, reclass_cat, grouped, nitrogen, source) %>% 
  rename(nitrogen_cat = nitrogen) %>% 
  filter(source == "calag")
  
}

results <- lapply(ag_files_list, fx_merge) %>% 
  setNames(c(2012, 2016, 2019))

# make into data frames

fx_df <- function(result){
  df_name <- data.frame(result)
}

dfs <- lapply(results, fx_df) %>% 
  setNames(c(2012, 2016, 2019))

# calculate stored carbon and nitrous oxide emissions for each pixel (900 sq m)

fx_ghg_calc <- function(dfs) {

dfs %>% 
  left_join(carbon_vals, by = "grouped") %>%
  mutate(mt_900 = (total_mt*.09)) %>% # MT carbon per hectare multiplied by .09 to get metric tons of carbon per pixel (900 sq m)
  left_join(lut_n, by = "nitrogen_cat") %>% 
  mutate(lbs_n_pixel = (n_rate_lbs_acre*.222395)) %>% # nitrogen application rate (pounds per acre) multiplied by .222395 to get pounds of N applied per per pixel
  mutate(emit_n_lbs_pix = (lbs_n_pixel * .0175)) %>% # 1% of nitrogen escapes at NO emissions
  dplyr::select(!c(n_rate_lbs_acre, lbs_n_pixel)) %>% 
  mutate(stock_abvgc_mtco2e_pixel = (mt_900*1)) %>%  # multiply metric tons of carbon by 3.67 to get MT of CO2 equivalent # Decided to report as MT instead, replace 3.67 value w 1 to not break rest of code
  mutate(emit_no_mtco2e_pix = emit_n_lbs_pix*298*0.000453592) # multiply pounds to NO emissions by 298 to convert to pounds CO2e, then by 0.000453592 to get metric tonnes 

}

ghgs <- lapply(dfs, fx_ghg_calc) %>% 
  setNames(c(2012, 2016, 2019))

ghg_dfs <- lapply(ghgs, fx_df) %>% 
  setNames(c(2012, 2016, 2019))

# add soil

fx_soil <- function(ghg_dfs) {
  
ghg_dfs %>%  
  merge(soil, by = "pointid") %>% 
  mutate(soil900 = (soc0_30*900)) %>% #per m^2 to per 900 m^2
  mutate(soilMT = (soil900/1000000)) %>%  # grams to metric tons of organic carbon
  mutate(stock_soilc_mtco2e_pix = soilMT*1) # convert to CO2e #Decided to report as MT instead, replace 3.67 value w 1 to not break rest of code
}

soil_results <- lapply(ghg_dfs, fx_soil) %>% 
  setNames(c(2012, 2016, 2019))

soil_results_dfs <- lapply(soil_results, fx_df) %>% 
  setNames(c(2012, 2016, 2019))

# name these just to preview
combined_2012_df <- soil_results_dfs[[1]]
combined_2016_df <- soil_results_dfs[[2]]
combined_2019_df <- soil_results_dfs[[3]]

# clean up results and summarize values
fx_all_clean <- function(df) {
df %>% 
  dplyr::select(reclass_cat, stock_abvgc_mtco2e_pixel, stock_soilc_mtco2e_pix, emit_no_mtco2e_pix) %>% 
  group_by(reclass_cat) %>% 
    summarise_all(.funs = c(sum="sum"), na.rm = TRUE) %>% 
    adorn_totals()
}

summaries <- lapply(soil_results_dfs, fx_all_clean) %>% 
  setNames(c(2012, 2016, 2019))

summaries_dfs <- lapply(summaries, fx_df) %>% 
  setNames(c(2012, 2016, 2019))

names <- names(summaries_dfs)

res_list <- vector("list", length = length(names)) %>% 
  setNames(names)

# lastly, make some tables

for(i in names){

loop_total_table <- soil_results_dfs[[i]] %>% 
  group_by(reclass_cat) %>% 
  summarize(pixels = n()) %>% 
  mutate(sqmeter = pixels*900) %>% 
  mutate(acreage = sqmeter/4047) %>% 
  merge(summaries_dfs[[i]], by = "reclass_cat") %>%
  filter(reclass_cat %in% c("Fallow", "Fodder", "Orchard", "Row Crop", "Pastureland", "Vineyard", "Greenhouse")) %>% 
  mutate(net = (stock_soilc_mtco2e_pix_sum + stock_abvgc_mtco2e_pixel_sum)) %>% 
  adorn_totals() %>% 
  mutate(year = i)
  
res_list[[i]] <- loop_total_table
  
  }

total_tables <- lapply(res_list, fx_df) %>% 
  setNames(c(2012, 2016, 2019))

# and name them as objects

ag_acreage_12 <- total_tables[[1]]
ag_acreage_16 <- total_tables[[2]]
ag_acreage_19 <- total_tables[[3]]

```
